---
title: "Homework25"
format: html
editor: visual
---

Setting up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Loading Data

```{r}
daphnia <- read.csv(here("data", "Daphniagrowth.csv"))
glimpse(daphnia)
```

Making parasite factor data

```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
```

Making a box and wisker plot to see variation in the response. 

Coord_flip() flibs the axis so that it is easier to read the group names

```{r}
ggplot(daphnia, aes(x=parasite, y=growth.rate))+
  geom_boxplot()+
  theme_bw()+
  coord_flip()

```

Estimating mean
control - 1.2
MB- .85
PP - 1.1
PR - 0.55
degrees of freedom - 3

Making the ANOVA

```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

Plotting the data

```{r}
autoplot(model_grow, smooth.colour = NA)
```
The data from the Q-Q test is okay despite looking weird
- i guess this is because there are not many data points that are off and they are all at the tails of the data. 

Anova and summary to make infrences from the summary. 

```{r}
anova(model_grow)
```
Is there an affect at all of the treatments?
- Yes there is because the p value is quite small allowing a rejection of the null hypothesis of there being no variance.

What are the effects?

```{r}
summary(model_grow)
```

Control is labelled as intercept
- Intercept is always the first level when they are organized alphabetically. 

How to get mean growth rates with dpplyr functions

```{r}
sumDat <- daphnia |> 
  group_by(parasite) |> 
  summarise(meanGR = mean(growth.rate))
```

How one way test of variance works:
- F is the statistic and is the ratio of between/within variance. 
- If F is high then there is a larger between group variance than within group variance
      - So a high F value indicates more signifigance in the data. 


















