---
title: "Homework26"
format: html
editor: visual
---

setting up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
library(multcomp)
```

importing data

```{r}
peas <- read.csv(here("Data", "s_r_pea_sections.csv"))
glimpse(peas)
```

Making data factor

```{r}
peas$Treatment <- as.factor(peas$Treatment)
```

Making a box and wisker plot to get a look at the data and make guesses at outcomes.

```{r}
ggplot(peas, aes(x=Treatment, y=pea_section_length_ou))+
  geom_boxplot()+
  geom_jitter()+
  coord_flip()+
  theme_bw()
```

Degrees of freedom -> 4 Estimating Means Control -> 71 2% sucrose -> 63.5 2% glucose -> 64 2% fructose -> 58.3 1% fructose 1% glucose -> 57.9

Making an ANOVA model

```{r}
model_peas <- lm(pea_section_length_ou ~ Treatment, data = peas)
```

Plotting the model to check for the assumptions of the model

```{r}
autoplot(model_peas, smooth.colour=NA)
```

The data does seem to be relatively normally distributed as all the points are along the line of fit even if they seem to trail off to the left there is a consistent pattern as to how they do.

Making an ANOVA to make inferences from the data.

```{r}
anova(model_peas)
```

Is there a significant difference? There is a significant difference between the control group and the treatment groups (F 4,45=49.368, DF=4, p=6.737e-16).

What are the effects?

```{r}
summary(model_peas)
```

The model shows that the variance in pea shoot length is almost completely explained by the variation between treatments (R-squared = 0.8144). The model shows that there is a significant difference between the 1% fructose 1% glucose treatment and the other treatments(p=2e-16). 

Doing a Tukey test between the different treatments.

```{r}
Tukey_peas <- glht(model_peas, linfct = mcp(Treatment = 'Tukey'))
summary(Tukey_peas)
```
According to the Tukey test there wasn't a significant difference between treatments between 2% glucose adn 2% fructose, 2% glucose and 1 percent fructose 1 percent glucose, and 2% fructose and 1 percent fructose 1 percent glucose (p=0.829091, p=0.725619, p=0.999688). All other treatment comparisons show a significant difference (p=1e-04).

This makes it easier to read the results
```{r}
cld(Tukey_peas)
```





