---
title: "Homework16"
format: html
editor: visual
---

set up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Importing data set into R and examing data

```{r}
cover <- read.csv(here("Data", "donnerville_percent_cover.csv"))
summary(cover)
glimpse(cover)
```

converting the data types of certain columns to factor data and date format

```{r}
cover$Site <- as.factor(cover$Site)
cover$Plot.type <- as.factor(cover$Plot.type)
cover$cover_type <- as.factor(cover$cover_type)
cover$Date <-  dmy(cover$Date)
```

Standardizing data by fixing data entry mistakes

```{r}
levels(cover$cover_type)
fix <- which(cover$cover_type %in% c("Vascular","vascular", "vasc "))
cover$cover_type[fix] <- "Vascular"
cover$cover_type <- factor(cover$cover_type) #releveling the factor
```

not sure why this is not standardizing I did it the same way from the notes that I got, since I was not in class and poked around on the internet but couldn't find anything helpful. vasc is still vasc but I don't get any sort of error message.

finding mean cover by plant type

```{r}
cover %>% summarise(mean(Percent)) %>% slice(1:12) %>% filter(cover, cover_type == Bare) #why does this line of code not work but putting the filter for cover_type first does?
cover %>% filter(cover_type == "Bare") %>% slice(1:3) %>% summarise(mean(Percent))
```

```{r}
cover %>% filter(cover_type == "Bryophytes") %>% slice(1:3) %>% summarise(mean(Percent))
```

```{r}
cover %>% filter(cover_type == "Vascular") %>% slice(1:4) %>% summarise(mean(Percent))
```

finding average vascular plant cover across all sites

```{r}
cover %>% filter(cover_type == "Vascular") %>% summarise(mean(Percent))
```
