---
title: "Biostats - Homework 20"
format: html
editor: visual
---

### Introduction

The objective of this assignment is to have you continue practicing the skills you have learned in Chapters 3 and 4 of Getting Started with R and reinforce the ideas we discussed in class regarding the Weissgerber et al. paper. Please complete all of the following steps:

1.  Write the code to set up your workspace.

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

2.  Read in the dataset called soil_data.csv and name it "soil."

```{r}
soil <- read.csv(here("Data", "soil_data.csv"))
```

3.  Create a histogram for the SOM_percent column in the dataframe. Accept the default number of bins.

```{r}
ggplot(soil, aes(x = SOM_percent))+
  geom_histogram()
```

How would you describe the mean_SOM data based on viewing the histogram? Write your answer below:

I would describe the mean SOM_percent as being unrepresentitive of the data since it has a large cluster around 5-15% but also a notable number of samples going up to around 55%.

4.  Create a new dataframe called SOM by calculating for each forest the mean soil organic matter and the standard deviation of soil organic matter. Make sure to name variables you create "mean_SOM" and "sd_SOM" so that the next code chunk I give you works.

```{r}
SOM <- soil |> group_by(
  Forest) |> summarize(
    mean_SOM = mean(SOM_percent),
    sd_SOM = sd(SOM_percent))

Nit <- soil |> group_by(
  Forest) |> summarise(
    mean_n = mean(Nitrate_N_ppm, na.rm= TRUE), 
    sd_n = sd(Nitrate_N_ppm, na.rm= TRUE))
```

5.  Now run the following code chunk which will create a bar plot of mean soil organic matter per forest, with standard deviation as error bars.

```{r}
ggplot(SOM, aes(Forest, mean_SOM))+
  geom_errorbar(aes(ymax = mean_SOM + sd_SOM, ymin = mean_SOM - sd_SOM), width = 0.2)+
  geom_col(fill = "white", color = "black")+
  ylab("Mean percent soil organic matter")+
  xlab("Forest name")
```

Based on what you see in the bar chart, how would you describe the pattern of SOM between the 4 different forests? Write your answer here:

I would describe it to be that when there is a higher mean percent soil organic matter there is a larger standard deviation between the data in the data set. This implies that there are not many samples showing higher percentages of soil organic matter.

Now create a boxplot of the SOM_percent data. Use Forest as the x-axis variable and overlay the point data on top of the boxplot. Make the points blue and somewhat transparent. Call the x-axis "Forest name" and the y-axis "Percent soil organic matter" (hint: See homework 18...). (x,y)

```{r}
ggplot(soil, aes(Forest, SOM_percent))+
  geom_boxplot()+
  geom_point(size = 2, color = 'blue', alpha = 0.1)+
  xlab("Forest name")+
  ylab("Percent soil organic matter")


```

Based on what you see in the box plot, write out your answers the following questions:

-   how would you describe the pattern of SOM between the 4 different forests?

    I would describe it as being a very tight standard deviation between the four forests except for the kip tract.

-   What do you notice that is different from how the data were displayed in the bar plot?

    It shows the mean within the IQR, or at least im pretty sure thats what the box is right? This shows which direction is having a larger skew on the data.

-   Which plot type is more informative? Explain your answer.

I think that the box plot is more informative since they show how the data is skewing the averages and helps to show that the higher amount of soil organic matter in the kip tract is a very wide range not a consistent difference from the other sites.
