---
title: "Homework21"
format: html
editor: visual
---

# My first Chi-square contingency analysis

setup

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

import data

```{r}
lady <- read.csv(here("Data", "ladybirds.csv"))
glimpse(lady)
```

Grouping data for analysis

```{r}
totals <- lady |>
  group_by(Habitat, colour) |>
  summarise(total_number = sum(number))
```

Making a bar chart

```{r}
ggplot(totals, aes(x = Habitat, y = total_number, fill = colour))+
  geom_bar(stat = 'identity', position = 'dodge')+
  scale_fill_manual(values = c(black = "black", red = "red"))
  
```

making a matrix of lady data to do a chi sq

```{r}
ladymat <- xtabs(number ~ Habitat + colour, data = lady)
# this does the same thing as the group_by()and summarise() functions
```

doing a chi-square test a test on the significance of your data.

```{r}
lady.chi <- chisq.test(ladymat) 
names(lady.chi)#gives more information on the chisq test
```

Read in the ozone data

```{r}
ozone <- read.csv(here("Data","ozone.csv"))
glimpse(ozone)
```

Making two histograms of the data to see if the means are different across groupings.

changing Garden.location and Garden.ID to factor data

```{r}
ozone$Garden.location <- as.factor(ozone$Garden.location)
ozone$Garden.ID <- as.factor(ozone$Garden.ID)
ozone$Ozone <- as.factor(ozone$Ozone)
ozone$Ozone
levels(ozone$Garden.ID)
levels(ozone$Garden.location)
levels(ozone$Ozone)
```

```{r}
ggplot(ozone, aes(ozone))+
  geom_histogram(binwidth = 10)+
  facet_wrap(~ Garden.location, ncol = 1)+
  theme_bw()
#Not sure what the error here is I've tried changing a few things
```

doing a t test

```{r}
t.test(ozone ~ Garden.location, data = ozone)
```

i think I might have messed up the data somehow while I was making garden location and ID factor data.
