---
title: "Homework25"
format: html
editor: visual
---

Setting up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Loading Data

```{r}
daphnia <- read.csv(here("data", "Daphniagrowth.csv"))
glimpse(daphnia)
```

Making parasite factor data

```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
```

Making a box and wisker plot to see variation in the response.

Coord_flip() flibs the axis so that it is easier to read the group names

```{r}
ggplot(daphnia, aes(x=parasite, y=growth.rate))+
  geom_boxplot()+
  theme_bw()+
  coord_flip()

```

Estimating means: control - 1.2 MB- .85 PP - 1.15 PR - 0.55 
degrees of freedom - 3
Error degrees of freedom are 36 from 40 samples minus the 4 treatments. 

Making the ANOVA
we can still use the lm() function since it is a linear model

```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

Plotting the data

```{r}
autoplot(model_grow)
```
Residuals vs Fitted:
The straight line shows that a linear model is a good choice for this data. 
Normal Q-Q:
The data from the Q-Q test is okay despite looking weird - i guess this is because there are not many data points that are off and they are all at the tails of the data.
Scale-Location:
Are the data homogeneous among the data set
Constant Leverage:
Are any points have a larger affect on the data. 

Anova and summary to make infrences from the summary.

```{r}
anova(model_grow)
```

Is there an affect at all of the treatments? - Yes there is because the p value is quite small allowing a rejection of the null hypothesis of there being no variance.
There is a statistically significant effect of parasite treatment on growth rate of daphnia (F(as a subscript 3, 36)=32.325, p=2.57e-10). 

What are the effects?

```{r}
summary(model_grow)
```
There is a significant difference between the Metschinikowia bicuspidata (p < 0.0001) and the control and 




Control is labelled as intercept - Intercept is always the first level when they are organized alphabetically.


How to get mean growth rates with dpplyr functions

```{r}
sumDat <- daphnia |> 
  group_by(parasite) |> 
  summarise(meanGR = mean(growth.rate))
```

How one way test of variance works: - F is the statistic and is the ratio of between/within variance. - If F is high then there is a larger between group variance than within group variance - So a high F value indicates more significance in the data.


How to plot the means difference
```{r}
ggplot(daphnia,aes(parasite, growth.rate, color=parasite))+
  geom_point()
```



