---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("Data", "CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- as.factor(crabs$Sex)
levels(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.

    ```{r}
    female_crabs <- crabs |> filter(Sex == "F")
    ```

2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:

    I decided that a regression was the best choice because the response variable, height, is ratio data and we are trying to see if there is a relationship between height and weight, the predictor variable.

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

```{r}
ggplot(female_crabs, aes(x=Height, y=Weight))+
  geom_point()+
  theme_bw()
```

remove an outlier
```{r}
female_crabs <- female_crabs |> filter(Height < 2) 
```


making a regression model

```{r}
model_fcrabs <- lm(Weight~Height, data=female_crabs)
```

visualizing the results

```{r}
autoplot(model_fcrabs, smooth.colour = NA)
```
These test the assumptions of a linear regression.
equally distributed(shown in top right plot) and equal variance in the data(shown in bottom left, we are looking for a lack of a pattern).




making an anova and summary table

```{r}
anova(model_fcrabs) 
  
summary(model_fcrabs)
 
```
The R-squared number is what percent of the crabs weight is determined by height. 
0.6715 means that 67 percent of the weight is determined solely by the height of the crabs. 
This is good since we are explaining 67% of the relationship with one variable. 
A perfect line of best fit means that every data point goes through the line and R-squared = 1. 



making regression lines to visualize results and to look for any relationship. 
```{r}
ggplot(female_crabs, aes(x=Height, y=Weight))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()
```

Accorinding to the results there is a signifigant positive relationship between Weight and Height in female crabs (Sumsq = 4.23, DF=1, F=613.566)

Not really sure which one is the test statistic. 









