---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("Data", "CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- as.factor(crabs$Sex)
levels(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.

    ```{r}
    female_crabs <- crabs |> filter(Sex == "F")
    ```

2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:

    I decided that a regression was the best choice because the response variable, height, is ratio data and we are trying to see if there is a relationship between height and weight, the predictor variable.

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

```{r}
ggplot(female_crabs, aes(x=Weight, y=Height))+
  geom_point()+
  theme_bw()
```

making a regression model

```{r}
model_fcrabs <- lm(Height~Weight, data=female_crabs)
```

visualizing the results

```{r}
autoplot(model_fcrabs, smooth.colour = NA)
```

making an anova and summary table

```{r}
anova(model_fcrabs) |>
  View()
summary(model_fcrabs) |>
  View()
```
making regression lines to visualize results and to look for any relationship. 
```{r}
ggplot(female_crabs, aes(x=Weight, y=Height))+
  geom_point()+
  geom_smooth(method = 'lm')+
  theme_bw()
```

Accorinding to the results there is a signifigant positive relationship between Weight and Height in female crabs (Sumsq = 4.23, DF=1, F=613.566)

Not really sure which one is the test statistic. 









